---
title: "Problem 3 - Asp Viper Management"
author: "Lindsey Pekurny"
date: "10/9/2020"
output: html_document
---

The Asp Viper, *Vipera aspis*, is a a threatened European snake. The data used in this analysis was gathered from 415 individuals amongst 6 populations of Asp Viper in the Jura Mountains of northern Switzerland. In general, snakes are very cryptic and can be difficult to study. Snakes are unfairly villainized in most cultures, however, they play critical roles in ecosystems, and thus deserve our protection. Using the data below, I want to determine population status - is this population growing or decreasing? Regardless of increasing or decreasing population sizes, I want to understand which age class most affects population growth. Based on these observations, I will create management recommendations.

Data:

```{r}

# Playing with the data

load("~/Quan/COMADRE_v.4.20.9.0.RData")

# Asp Viper
Vipera <- grep(pattern = 'Vipera',x=comadre$metadata$Genus)
head(Vipera)

# Isolate matrix for Asp Viper

vipmat <-  comadre$mat[2231][[1]]
vipmat <- vipmat$matA          ### Using this matrix


# Information about the study

comadre$metadata$StudyDuration[2231] ## study took 17 years
comadre$metadata$Authors[2231]
comadre$metadata$YearPublication[2231]



```
The matrix provided in this data set has broken the asp vipers into 5age classes determined by size. Age is dependent upon size, and so juveniles were classified as individuals less than 37cm in length. Sub-adults were broken into 3 groups, first years (30-37cm), second years (37-44cm), and third years (44-50cm). Adults were any snakes larger than 50cm.


Analysis of matrix:

``` {r}

library(popbio)

eigen.analysis(vipmat, zero = FALSE)

```

From the eigen analysis, I know the dominant eigenvalue is 0.88. The population is therefore decreasing by 12% each year. In order for the population to remain stable, 30% of the population needs to be juveniles and 30% of the population needs to be adults. Since these two age classes need the majority of the population for the overall population to remain stable, management for this species should prioritize the protection of these groups. Of these two groups, however, adults have the highest reproductive value (7.01). Juveniles have a reproductive value of 1. Based on this knowledge, the most conservation effort should be placed on the adults in this population.

Juveniles (0.66) has the highest level of sensitivity, followed by adults (0.62) - they are both very similar. If a small change in the number of individuals within this group occurs, the overall population will be effected. Instead of making a decision based on sensitivty alone, I used the elasticity for this data. Elasticity is proportional sensivity, whereas sensitivity is absolute change. Therefore, using elasticity gives us an understanding of population dynamics in proportions rather than absolute values. When looking at the elasticity for each age group the adult population is the most elastic, indicating if the adult population increased by just 1%, the population growth rate will go up by 0.53%. 

Using this knowledge, I would recommend conservation efforts prioritize the adult age group. More data should be gathered on preferred adult habitat type, and this habitat should be preserved. The creation of additional habitat would also benefit these snakes. Unfortunately, snakes are often persecuted by humans. Education iniatives could help people understand the importance of snakes in our ecosystems and prevent further persecution.

